---
title: "Reproducible Research: Peer Assessment 1"
author: "Haichen"
date: "3/15/2021"
output: html_document
keep_md: true
---

This assignment will be described in multiple parts. You will need to write a report that answers the questions detailed below. Ultimately, you will need to complete the entire assignment in a single R markdown document that can be processed by knitr and be transformed into an HTML file.

Throughout your report make sure you always include the code that you used to generate the output you present. When writing code chunks in the R markdown document, always use echo = TRUE so that someone else will be able to read the code. This assignment will be evaluated via peer assessment so it is essential that your peer evaluators be able to review the code for your analysis.

For the plotting aspects of this assignment, feel free to use any plotting system in R (i.e., base, lattice, ggplot2)

Fork/clone the GitHub repository created for this assignment. You will submit this assignment by pushing your completed files into your forked repository on GitHub. The assignment submission will consist of the URL to your GitHub repository and the SHA-1 commit ID for your repository state.

NOTE: The GitHub repository also contains the dataset for the assignment so you do not have to download the data separately.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Coursera/Reproducible Research/project 1/RepData_PeerAssessment1")
```

## Loading and preprocessing the data

```{r Get Data}
if (!file.exists('activity.csv')) {
  unzip(zipfile = "activity.zip")
}

activityData <- read.csv(file="activity.csv", header=TRUE)
```


## What is mean total number of steps taken per day?

```{r Total steps per day Histogram}

totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)

hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")

# Calculate and report the mean and median of total steps taken per day
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
```

```{r Mean and Median}
# Calculate and report the mean and median of total steps taken per day
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
meanSteps
medSteps
```


## What is the average daily activity pattern?


```{r Pattern plot}
# Make a time-series plot of the 5-minute interval 
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
# Which 5-minute interval across all days contain the maximum number of steps
maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]
```

## Imputing missing values


```{r Imputing Missing Values}

missingVals <- is.na(activityData$steps)

imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, 
                                                                        meanStepsByInt$interval)],
                                             activityData$steps))

# Make a histogram of the total number of steps taken each day and report the mean and median.
impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
```

```{r Diff in Mean adn Med}
impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)
diffMean
diffMed 
diffTotal
```

## Are there differences in activity patterns between weekdays and weekends?


```{r Diff pattern between weekdays and weekends}
# Calculate and report the mean and median of total steps taken per day
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
      return ("weekend")
  else
      stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)

# Make a panel plot containnig a time-series plot of the 5-minute interval
# and the average number of steps taken across all weekdays or weekends
meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```